### 2016-10-01 paper_intersections.tex
Theoretical FLOPS counts.

%% 
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%% SP 2008/03/01

%\documentclass[preprint,12pt]{elsarticle}
 \documentclass{elsarticle}
%\documentclass[3p,12pt,authoryear]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

\usepackage{hyperref}
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = red %Colour of citations
}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}
\usepackage{subfig}

% alphabetical enumeration
\usepackage{enumitem}

%tables
\usepackage{booktabs}
\usepackage[vlined, linesnumbered, ruled]{algorithm2e}
%\usepackage{float}
%\newfloat{algorithm}{t}{lop}
\usepackage{array}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{esint}

\newtheorem{theorem}{Theorem}[section]
%\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
%\newtheorem{remark}[theorem]{Remark}
\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{thmproblem}{Problem}
%\newtheorem{exercise}[theorem]{Cvičení}

%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

% just for our notes
\usepackage[usenames,dvipsnames]{xcolor}   %colors

\journal{Applied Mathematics and Computation}

%commands:
%\newcommand{\defref}[1]{\hyperref[#1]{Def.~\ref{#1}}}
\newcommand{\prob}[1]{Problem~{#1}}
\newcommand{\fig}[1]{\hyperref[#1]{Figure \ref{#1}}}
\newcommand{\figpath}{figures/}

%math:
\def\vc#1{\mathbf{\boldsymbol{#1}}}     % vector
\def\abs#1{\left|#1\right|}
\def\avg#1{\langle#1\rangle}
\def\d{\mathrm{d}}
\def\norm#1{\| #1 \|}
\def\abs#1{| #1 |}
\def\prtl{\partial}
\newcommand{\dd}{\; \mathrm{d}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\bx}{\vc{x}}
\newcommand*\rfrac[2]{{}^{#1}\!/_{#2}}

\newcommand{\noteJB}[1]{{\color{Blue} \textbf{JB: } \textit{#1}}}
\newcommand{\notePE}[1]{{\color{Orange} \textbf{PE: } \textit{#1}}}
\newcommand{\plucker}{Pl\"{u}cker }
\newcommand{\nface}{$n$-face }
\newcommand{\nfaces}{$n$-faces }
\newcommand{\ngh}{NGH }

\newdefinition{mdef}{Definition}%[section]

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

%\title{Algorithm for Intersections of Nonmatching Grids of Different Dimensions}
\title{Fast Algorithms for Intersection of Non-matching Grids Using \plucker Coordinates.}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author[adr]{Jan B{\v r}ezina\corref{cor1}}
\ead{jan.brezina@tul.cz}
\cortext[cor1]{Corresponding author.}

\author[adr]{Pavel Exner}
\ead{pavel.exner@tul.cz}
%\ead[url]{https://github.com/Paulie14/xfem\_project}

\address[adr]{Technical University of Liberec, Studentsk{\' a} 1402/2, 461 17 Liberec 1, Czech Republic}

\begin{abstract}
The XFEM and Mortar methods can be used in combination with non-matching or non-conforming grids 
to deal with problems on complex geometries. However the information about the mesh intersection must be provided.
We present algorithms for intersections between 1d and 2d unstructured multi component meshes 
and their intersections with a background unstructured 3d mesh. A common algorithm 
based on the advancing front technique is used for the efficient selection of candidate pairs among 
simplicial elements. Bounding interval hierarchy (BIH) of axes aligned bounding boxes (AABB) of elements
is used to initialize the front tracking algorithm. The family of element intersection algorithms
is built upon a line-triangle intersection algorithm based on the \plucker
coordinates. These algorithms combined with the advancing front technique can reuse the results of calculations 
performed on the neigbouring elements and reduce the number of arithmetic operations. Barycentric coordinates
on each of the intersecting elements are provided for every intersection point. Benchmarks 
of the element intersection algorithms are presented and three variants of the global intersection 
algorithm are compared on the meshes raising from a hydrogeological application.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
non-matching grid \sep
non-conforming mesh \sep
mesh intersection \sep
mixed-dimensional mesh \sep
\plucker coordinates \sep 
advancing front method

%% PACS codes here, in the form: \PACS code \sep code
%\PACS 02.60.Lj \sep        %Ordinary and partial differential equations; boundary value problems
%\PACS 02.60.Jh             %Numerical differentiation and integration

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
%\MSC[2010] 65N30 \sep %    Finite elements, Rayleigh-Ritz and Galerkin methods, finite methods
%\MSC[2010] 35J05  %    Laplacian operator, reduced wave equation (Helmholtz equation), Poisson equation

\end{keyword}

\end{frontmatter}

%% \linenumbers

%% main text
\section{Introduction}
The grid intersection algorithms are crucial for several techniques that try to overcome some limitations of the classical finite element method.
The Chimera method \cite{brezzi_analysis_2001}, also called overset grid, and similar Nitche method \cite{massing_efficient_2013}
allow solution of the problems with changing geometry as in the fluid-structure problems. 
The Mortar method \cite{belgacem_mortar_1999} allows domain decomposition, independent meshing of domains, and supports sliding boundaries.
However our primal motivation is usage of XFEM methods \cite{fries_extended/generalized_2010} and non-matching meshes of mixed dimension in groundwater models.

The realistic models of groundwater processes including the transport processes and geomechanics have to deal with 
a complex nature of geological formations including fractures and wells. Although of a small scale, these features may have significant impact 
on the global behavior of the system and their representation in the numerical model is imperative.
One possible approach is to model fractures and wells as lower dimensional objects and introduce their coupling with the surrounding continuum. 
The discretization then leads to the meshes of mixed dimensions, i.e. composed of elements of different dimension. This approach 
called mixed-dimensional analysis in the mechanics \cite{bournival_mesh-geometry_2008} is also studied in the groundwater context, see e.g. 
\cite{martin_modeling_2005}, \cite{fumagalli_numerical_2011}, \cite{brezina_analysis_2015} and 
already adopted by some groundwater simulation software, e.g FeFlow \cite{trefry_feflow:_2007} and Flow123d \cite{flow123d}.
Nevertheless as the complexity of the geometry increase (e.g. when lots of fractures are randomly generated) the compatible meshing becomes painful or even 
impossible. In order to avoid these difficulties we may discretize the continuum and every fracture and well independently, getting a non-matching 
(or incompatible) mesh of mixed dimensions and then apply XFEM to represent jumps of the solution on the fractures or singularities 
around the wells. The prerequisite for such approach is a fast and robust algorithm for calculating intersections of individual meshes.

We consider a composed mesh $\mathcal T$ consisting of simplicial meshes $\mathcal T_i$ of dimensions $d_i \in \{1,2,3\}$, $i=1,\dots,N_\mathcal{T}$
in the 3d ambient space. We assume that every mesh $\mathcal T_i$ is a connected set with no self intersection.
Further we assume only single 3d mesh $\mathcal{T}_1$.
The mesh intersection problem is to find all pairs of elements $L\in \mathcal{T}_i$, $K\in \mathcal{T}_j$, $i\ne j$ that have non-empty intersection 
and to compute that intersection. The mesh intersection problem consists of the two parts: First, generate a set of candidate pairs $(K, L)$. 
Second, compute the intersection of a particular pair.

According to our knowledge, there are lots of works using non-matching grids, yet only few of them discuss algorithms how to compute their intersections. 
Gander and Japhet \cite{gander_algorithm_2013} present the PANG algorithm for 2d-2d and 3d-3d intersections that can be used e.g. for mesh overlapping methods.
They use the advancing front technique to get candidate pairs in linear time. The algorithm is part of the DUNE library \cite{bastian_towards_2005}.
Massing, Larson, and Logg \cite{massing_efficient_2013} present an algorithm for 2d-3d intersections as part of their implementation of the Nitche method 
which is part of the Doplhin project \cite{Dolphin}. They use axes aligned bounding boxes of elements (AABB) and bounding interval hierarchy (BIH) 
to get intersection candidate pairs of elements, the GTS library \cite{gtslib} is used for 2d-3d intersections. 
Finally, there is the work of A. Elsheikh and M. Elsheikh \cite{elsheikh_reliable_2012} presenting an algorithm for 2d-2d mesh union operation which includes 
calculation and imprinting of the intersection curves. They exploit binary space partitioning for search of initial intersection and 
the advancing front method for intersection curve tracking.

In this paper we present a new approach to intersection calculation of simplicial elements of different dimensions based on the \plucker coordinates,
further developing the algorithm of Platis and Theoharis \cite{platis_fast_2003} for ray-tetrahedron intersections. Element intersections based on 
\plucker coordinates are combined with the advancing front method which allows us to reuse \plucker coordinates and their products between neigbouring
elements and reduce the number of arithmetic operations.

The paper is organized as follows. In Section \ref{sec:element intersecitons} the algorithms for
1d-2d, 1d-3d and 2d-3d intersections of simplices are described. In Section \ref{sec:front_advancing} we discuss our implementation of the advancing front 
technique and usage of AABB and BIH for its initialization. Finally, in Section \ref{sec:benchmarks}, we present benchmarks and comparison of individual algorithms.

\section{Element Intersections}
\label{sec:element intersecitons}
In this section, we present algorithms for computing intersection of a pair of simplicial elements of a~different dimension in the 3d ambient space. 
In particular we address intersection algorithms for 1d-2d, 1d-3d, 2d-3d pairs of elements. We have implemented the case 2d-2d as well, however the treatment 
of the special cases is quite technical and not fully completed yet. The fundamental idea is to compute intersection of 1d-2d simplices using the 
\plucker coordinates and reduce all other cases to this one.

We denote $S_i$ a simplicial element with $i+1$ vertices (of dimension $i$). We call vertices, edges, faces and simplices itself the \nfaces and we denote
$M_i$ the set of all \nfaces of the simplex $S_i$.
In general, an intersection can be a point, a line segment or a polygon called \emph{intersection polygon} (IP) in common.
The intersection polygon is represented as a list of its corners called \emph{intersection corners} (IC). The IP data structure keeps also 
reference to the intersecting simplices. A data structure of a single IC consists of:
\begin{itemize}
 \item the barycentric coordinate $\vc w_K$ of IC on $K$,
 \item the dimension $d_K$ of the most specific \nface the IC lies on,
 \item the local index $i_K$ of that \nface on $K$,
\end{itemize}
for each intersecting element $K$ of the pair. The pair $\tau_K = (d_K, i_K)$
is called the topological position of the IC on $K$.

\subsection{\plucker Coordinates}
\plucker coordinates represent a line in 3d space.
Considering a line $p$, given by a point $\vc A$ and its directional vector $\vc{u}$, 
the \plucker coordinates of $p$ are defined as
\[ \pi_p = (\vc{u}_p, \vc{v}_p) = (\vc{u}_p, \vc{u}_p\times \vc A). \]
Further we use a permuted inner product
\[\pi_p \odot \pi_q = \vc{u}_p\cdot \vc{v}_q + \vc{u}_q \cdot \vc{v}_p. \]

The sign of the permuted inner product gives us the relative position of the two lines, 
see \fig{fig:plucker_products}.

\begin{figure}[!htb]
  \begin{center}        
    \includegraphics[width=0.7\textwidth]{\figpath plucker_product.pdf}
%    \includegraphics{\figpath plucker_products.pdf}
  \end{center}
  \caption{Sign of the permuted inner product is related to the relative position of the two oriented lines. Dashed line symbolizes that the line is in the back, the lines intersect in the middle case.
           \notePE{circle dot permuted inner product}}
  \label{fig:plucker_products}
\end{figure}

%Notice the condition on the orientation of the triangle sides. To always satisfy it, we use a
%reference simplex, in which the numbering of nodes and sides and also the orientation is fixed.
%The barycentric coordinates of the intersection corner can be computed directly from the \plucker products
%and then we can easily obtain its real coordinates (see \cite{fris_dp_2015} for derivations and details).

\subsection{Intersection Line-Triangle (1d-2d)}
\label{sec:1d-2d}
\notePE{I would stick to chosen notation - triangle $S_2$, and possibly renamed side $s_i$...}
Let us consider a line segment $p$ with parametric equation 
\begin{equation}
    \label{eq:line_parametric}
    \vc X = \vc A + t\vc u,\ t\in (0,1)
\end{equation}
and a triangle $T$ given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. 
\begin{lemma}
The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same non-zero sign if and only if there
is an intersection point $\vc X$ on the $p$ and inside the triangle $T$. 
The barycentric coordinates of $\vc X$ on $T$ are
\begin{equation}
  \label{eq:bary_centric}
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }.
\end{equation}
\end{lemma}
\begin{proof}
Using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ has \plucker coordinates $(\vc u, \vc u \times \vc X)$ since these are invariant to change of the initial point. 
Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
The point $\vc X$ is inside of $T$ if and only if $w_i>0$ for all $i=0,1,2$.
\end{proof}
Having the barycentric coordinates of $\vc X$ on $T$, we can compute also its local coordinate on $p$ from its parametric form:
\begin{equation}
   \label{eq:line}
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\end{equation}
We use $i$ with maximal $|u_i|$ for practical computation.

The calculation of the intersection proceeds as follows:
\begin{enumerate}
    \item Compute or reuse \plucker coordinates and permuted inner products: $\pi_p$, $\pi_i$, $\pi_s \odot \pi_i$, for $i=1,2,3$.
    \item Compute barycentric coordinates $w_i,\ i=1,2,3$ using \eqref{eq:bary_centric}.
    \item If any $w_i$ is less than $\epsilon$, there is no intersection, return empty IP.  
    \item If all $w_i$ are greater than $\epsilon$, we set $\tau_T = (2, 0)$ for the IC.
    \item If one $w_i$ is less than $\epsilon$, intersection is on the edge $s_i$, we set $\tau_T =(1,i)$.
    \item If two $w_i$ are less than $\epsilon$, intersection is at the vertex $V_i$, we set $\tau_T=(0,i)$.
    \item \label{item:coplanar} If all $w_i$ are less than $\epsilon$, the line is coplanar with the triangle, both objects are 
        projected to the plane $x_i=0$ where $i$ is the index of maximal component of the triangle's normal vector.
        Every pair $p$, $s_i$ is checked for an intersection on $T$ boundary either inside $s_i$ or at a vertex $V_i$, setting the topological info $\tau_T$ to 
        $(1, i)$ or $(0, i)$, respectively. At most two ICs are obtained.
    \item For each IC the barycentric coordinates $(1-t, t)$ on the line $p$ are computed according to \eqref{eq:line}. 
    \item If $t\in (-\epsilon, \epsilon)$ or $t\in (1-\epsilon, 1+\epsilon)$, we set $\tau_p = (0,0)$ or $\tau_p = (0,1)$, respectively.
    \item If $t\notin (-\epsilon, 1+\epsilon)$, the IC is eliminated.
\end{enumerate}

We shall call every IC, that was created in the coplanar case, \emph{degenerate} since these will later
special treatment.

Checking the same sign of the inner products can be viewed as a geometric predicate for the presence of the intersection and orientation of the 
line with respect to the triangle. Adaptive-precision evaluation of the geometric predicates was designed by Schewchuk \cite{shewchuk_adaptive_1997}
and used for 2d-2d mesh intersections in \cite{elsheikh_reliable_2012}. However, we rather apply a fixed tolerance check for the zero 
barycentric coordinates and consistently keep the topological positions in this and related algorithms. \noteJB{Can we make the algorithm 
\emph{parsimonious} in the spirit of the Fortune \cite{fortune_stable_1989} quoted by Schewchuk? 
Seems that our problem is more local than the line example that was proven to be NP-hard.}

The algorithms for 1d-3d and 2d-3d intersections use simpler version of the 1d-2d intersection algorithm, in particular the search for ICs in the coplanar case 
(item \ref{item:coplanar}) is not necessary, and the test in the last point is not performed.
%and degenerate cases higher dimensional cases.

%\subsection{Intersection Triangle-Tetrahedron (2d-2d)}
%\noteJB{Just note that this algorithm can also be implemented using 1d-2d, but quite complex to not yet fully tested}

\subsection{Intersection Line-Tetrahedron (1d-3d)}
\notePE{degenerate not defined yet.. do you call it coplanar now?}
In this section we consider an intersection of a line segment $p$ given by the parametric equation \eqref{eq:line_parametric} with a tetrahedron 
$S_3$. The used algorithm is based on the 1d-2d algorithm and closely follows \cite{platis_fast_2003}. Our modification takes into account
intersection with the line segment and consistently propagates topological position of ICs.

\begin{algorithm}
  \caption{1d-3d intersection}
  \label{algo:13d}

  \DontPrintSemicolon
  \SetKw{and}{and}
  \SetKw{continue}{continue}
  \SetKw{break}{break}
  \KwIn{Tetrahedron $S_3$, line segment $p$.}
  \KwOut{List $I$ of ICs sorted along $p$.}  
  $I=\{\}$\; 
  \For{unmarked face $f$ of $S_3$}{
    $L$ = intersection($p,\ f$)\; \label{line:13-intersection}
    \lIf{ $L$ is none or degenerate}{\continue}
    \If{ $L$ is inside the edge $e$}{
        set $\tau_{S_3} = (1,e)$\;
        mark faces coincident with $e$
    }
    \ElseIf{ $L$ is at the vertex $v$}{
        set $\tau_{S_3} = (0,v)$\;
        mark faces coincident with $v$
    }
    append $L$ to $I$\; \label{line:13-append}
    \lIf{$\abs{I}=2$}{\break}
  }
  \lIf{$\abs{I}=1$ \and   $I$ is outside of $p$}{ erase $I$} \label{line:trimming}
  \ElseIf{$\abs{I}=2$}{    
    trim intersection with respect to the line segment $p$
  }
\end{algorithm}
Algorithm \ref{algo:13d} first computes line-face intersections for every face of $S_3$.
Tetrahedron has six edges, so 7 \plucker coordinates and 6 permuted inner products are computed at most. Precomputed coordinates and products are passed into the 1d-2d algorithm which is performed for the whole line $p$ (line \ref{line:13-intersection}).
If no IC is found, or coplanar case occurs in line-face computation, we continue to the next face. Note, ICs that would be created in coplanar case are to be found as ICs with the other faces, since they lie on edges.
Next, IC can be on an edge or at a vertex; then we set the correct topological position and mark the adjacent faces
to be skipped, since there cannot be another IC (and coplanar case has been checked already).
Finally at the line \ref{line:13-append}, we append the IC to the result and check whether the maximal amount of 
ICs has been reached.

After collecting line-tetrahedron ICs,
we do the line segment trimming from the line \ref{line:trimming} further. 
If we have only one IC, we check that it actually lies inside the line segment (otherwise, we throw it away).
If we have two ICs, and if both lie outside the line segment $p$, we eliminate both of them. 
If one of the ICs lies out of $p$,
we use the closest end point of the line segment instead and interpolate barycentric coordinates of the IC on $S_3$. The topological positions are updated as well.
The result of the algorithm is 0, 1, or 2 ICs, sorted by the parameter $t$ in the direction of the line $p$.

\subsection{Intersection Triangle-Tetrahedron (2d-3d)}
The intersection of a triangle $S_2$ and a tetrahedron $S_3$ is an $n$-side intersection polygon (IP), $n\le 7$. The sides of the polygon
lie either on sides of $S_2$ or on faces of $S_3$. Thus each vertex (IC) of the polygon
can arise either from side-face intersection, or from edge-triangle intersection, or be a vertex of $S_2$.
To get all ICs, we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner products, one for every side-edge pair.
Computation of IP consists of three stages: calculation of side-tetrahedron ICs (Section \ref{sec:sides}), 
calculation of edge-triangle ICs (Section~\ref{sec:edges}), ordering of ICs (Section~\ref{sec:ordering}).
The intersection corners are appended to the list $I$ as they are computed, however their order on the polygon boundary is 
defined by the \emph{connection tables} $F_g(:)$ and $F_p(:)$. Every side of the polygon that lies on \nface $x\in M_2\cup M_3$ is followed by an IC given by $F_g[x]$ 
and every IC $p$ is followed by the side that lies on $F_p[p]\in M_2\cup M_3$ (see \noteJB{possible figure}). The vertices of the polygon are ordered in the same 
way as is the order and orientation of the sides of $S_2$, that is counterclockwise around the interior with normal pointing to us.
There are special cases when the IP is degenerated to a line or a point. 
\notePE{next sentence: how should I understand 'may possibly be'; 'an IC' instead of 'some IC'}
If some IC may possibly be part of the such degenerated IP we put it into the list $J$.

\subsubsection{Intersections on sides of $S_2$}
\label{sec:sides}

\begin{algorithm}
  \caption{2d-3d intersection, ICs on sides of $S_2$}
  \label{algo:colect_23_ip_triangle}

  \DontPrintSemicolon
  \SetKw{continue}{continue}
  \KwIn{input data}
  \KwOut{List of ICs on sorted output data}  
  $F_g(:)=-1,\ F_p(:)=-1$\tcp*{Unset links.} 
  \For{side $s$ of $S_2$}{ \nllabel{line:one side loop}
    $L$ = intersection($s, S_3$)\;
    \lIf{ $\abs{L} = 0$}{\continue}
    \lIf{ $\abs{L} = 1$}{
        append $p$ to $J$
        \continue
    }
    \For{$p$ in $L$}{
%        \lIf{$p \ne I[-1]$ }{append $p$ to $I$}
        $p$ lies on $x\in M_3$\; \nllabel{line:m4_object}
        \lIf{$p$ lies at the vertex $v$ of $S_3$}{$x=v$}
        \uIf{$p$ is the first in $L$}{
            $F_g[x] = p$,\ $F_p[p] = s$\;
        }
        \Else($p$ is the last in $L$){
            $F_g[s] = p,\ F_p[p] = x$\;
            \lIf{$F_g[x] = -1$}{$F_g[x] = p$} \nllabel{line:back_link43}
        }
    }
  }  
%  \If{$I[-1] = I[0]$}{
%    $x = F_p[I[-1]]$\;
%    \If{$F_g[x]]=I[-1]$}{$F_g[x]=I[0]$}
%    $F_g[s_2] = I[0]$\;
%    remove $I[-1]$
%  }
\end{algorithm}

Algorithm \ref{algo:colect_23_ip_triangle} computes all ICs on the boundary of $S_2$.  It passes through every side $s$ of the triangle $S_2$ 
and computes the line-tetrahedron intersection $L$.
If the side is just touching $S_3$, we get $L$ with a single IC. These ICs will be rediscovered again in Algorithm \ref{algo:collect_23_ip_edges} 
with better topological information, however this is not the case if the touched edge $e$ of $S_3$ is coplanar with $S_2$ and the IC is inside of $e$.
To this end we save the IC into separate list $J$ and skip filling of the connection tables.
\notePE{here it is weird... there is no line appending the IC into resulting list.. Which is the resulting list??
I find the reference to 'line:one side loop' also incorrect..}
In the regular case, we process each of the two ICs in $L$ (loop on line \ref{line:one side loop}). The IC $p$ is added to the list $I$,
vertices of $S_2$ added twice are merged at final stage. Then (line \ref{line:m4_object}) we identify the \nface $x\in M_3$ 
the point $p$ lies on and set the tables $F_g$, $F_p$. For the IC at the vertex of $S_2$ we set $x$ to that vertex regardless of its position on $S_3$.
The backward temporary link on the line \ref{line:back_link43} is used later in Algorithm \ref{algo:collect_23_ip_edges} to allow proper continuation 
of the IC if it lies on edge or at vertex of $S_3$. The condition at the same line deals with the case that two sides of $S_3$ intersect 
the same object $x\in M_3$, in such case the backlink is unnecessary and would overwrite the correct link set by previous IC.

\subsubsection{Intersections on edges of $S_3$}
\label{sec:edges}
\begin{algorithm}  
  \caption{2d-3d intersection, ICs on edges of $S_3$}
  \label{algo:collect_23_ip_edges}

  \SetKw{And}{And}
  \SetKw{edgefaces}{edge faces}
  \SetKw{vertexfaces}{vertex faces}
  \DontPrintSemicolon
  \KwIn{$I$ with ICs on $S_2$ boundary, partially filled $F$}
  \KwOut{all ICs in $I$, complete $F$}

  \lFor{edge $e$ of $S_3$}{$L[e]$ = intersection($e, S_2$)}  \nllabel{line:12edges}
  \For{edge $e$ of $S_3$ with non-empty $L[e]$}{
    $p=L[e]$\;
    \uIf{ $p$ is inside $e$}{ 
        \nllabel{line:edge_faces}        
        $(f_0,\ f_1)$ = \edgefaces($e$)
    } 
    \Else($p$ at the vertex $v$ of $S_3$){       
            $(f_0,\ f_1)$ = \vertexfaces($v$,$L$)   \tcp*{Algorithm \ref{algo:vertex_faces}}
            \nllabel{line:mark_edges}
    }    
    \uIf{ $p$ is on the boundary of $S_2$ }{
         $p$ lies on edge or at vertex $x \in M_3$\;
         $q=F_g[x]$\tcp*{$q$ is already computed $p$}
         \lIf{$F_p[q]=x$}{$F_p[q]=f_1$} \nllabel{line:successor}
         \lElse{$F_g[f_0]=q$}  \nllabel{line:predecessor}
         $F_g[x]=-1$\tcp*{remove the backlink}
    }
    \Else{
         append $p$ to $I$\; 
         $F_g[f_0]=p,\ F_p[p]=f_1$\tcp*{overwrite the backlink} \nllabel{line:insideS2}
    }   
  }  
\end{algorithm}

\begin{figure}%[!htb]
  \begin{center}        
    \includegraphics[width=0.3\textwidth]{\figpath edge_faces.pdf}
  \end{center}
  \caption{Order of faces adjacent to the oriented edge $e$ pointing up.}
  \label{fig:edge_faces}
\end{figure}

Algorithm \ref{algo:collect_23_ip_edges} uses the line-triangle intersection algorithm for the edges of $S_3$ (line \ref{line:12edges}).
First, the intersection $L[e]$ is evaluated for every edge $e$. The loop produces ICs in the interior of $S_2$ and possibly those ICs with special position at the vertex or on the edge of $S_3$ already computed 
in Algorithm \ref{algo:colect_23_ip_triangle}. Then we pass through once again skipping the edges with none or degenerate $L[e]$.
For every IC $p=L[e]$ we first get its (generalized) faces that would appear before and after the IC on IP, assuming that $p$ is inside $S_2$.

The function \emph{edge faces} returns the adjacent faces $f_0$, $f_1$ to the edge $e$, on which the IC lies (see the situation in \fig{fig:edge_faces}). 
The faces are sorted using the sign of the permuted inner product in 1d-2d intersection. The order of faces matches 
the order of sides of IP if the sign is negative. If the sign is positive the function \emph{edge faces} returns face pair $(f_1, f_0)$.
If the IC is at the vertex $v$ of $S_3$, the function \emph{vertex faces} described later (Algorithm \ref{algo:vertex_faces}) is used.
It returns a pair of generalized faces (face or edge) adjacent to the IC $L[e]$ at the vertex $v$ of $S_3$. 
If IC $p$ is inside $S_2$, we add it into $I$ and set the connection tables (line \ref{line:insideS2}). 
However,
if $p$ is on boundary of $S_2$ it is already in $I$. Denoting $x$ the \nface of $S_3$ the IC lies on, we use the back-link from $x$ to 
IC $q\in I$ and if $q$ points back to $x$ we set the new successor $f_1$ to it (line \ref{line:successor}). 
Otherwise $f_0$ is used as the predecessor of $q$ (line \ref{line:predecessor}).

\subsubsection{Vertex Faces Algorithm}

\begin{algorithm}
  \caption{2d-3d intersection, vertex faces}  
  \label{algo:vertex_faces}

  \SetKw{return}{return}
  \SetKw{edgefaces}{edge faces}
  \DontPrintSemicolon
  \KwIn{vertex $v$ of $S_3$, $L[:]$ intersection results for edges of $S_3$}
  \KwOut{$(x_1, x_2)$, $x_1, x_2 \in M_3$, coincident with $v$ and intersected by the plane of $S_2$}
  $e_0,\ e_1,\ e_2$ edges coincident with $v$ oriented out of $v$
  $s[i] = L[e_i]$, for $i=0,1,2$, \;

  \uIf{$s[:]$ have 1 non-degenerate edge $e$}{ \nllabel{line:two_degenerated}
    \return pair of degenerate edges sorted according to \edgefaces($e$)    
  }

  \uElseIf{ $s$ have 1 degenerate edge $e$}{\nllabel{line:one_degenerated}
    $f$ is face opposite to $e$\;
    \uIf{other two edges $e_a$, $e_b$ have different sign}{
        $z=\edgefaces(e_a)$\;
        replace $g\in z$, $g\ne f$ with $e$, \return $z$
    }\lElse{
        append IC of $v$ to $J$, \return anything
    }
  }  
  \uElseIf{ $s$ have edge $e$ with sign opposite to other two}{ \nllabel{line:different_sign}
    \return edge faces($e$)\;
  }

  \Else($s$ have all signs same){ \nllabel{line:same_signs}
    append IC of $v$ to $J$, \return anything\;
  }  
\end{algorithm}

This function gets as a parameter IC $p$ at the vertex $v$ of $S_3$ which is a special case of anon-degenerate edge-triangle intersection. 
We use signs and degenerate indicator of ICs of the three edges coincident with $v$ to return generalized faces of $S_3$ 
preceding and succeeding $p$ on the polygons boundary assuming $p$ is at interior of $S_2$.
Possible cases are:
\begin{itemize}
 \item {\bf Single degenerated IC} (line \ref{line:one_degenerated}){\bf .} Let us denote $e$ the edge with degenerated IC 
 and $f$ the face between the other two edges.
 The other two (non-degenerates) edges may have either the opposite sign (the plane is cutting $S_3$) 
 or the same sign (the plane is touching $S_3$ at the edge $e$).
 In the first case, the call of edge faces for $e$ returns $(f_x,f)$ or $(f,f_x)$, then the vertex faces function returns $(e, f)$ or $(f, e)$, respectively.
 In the second case, we just return anything since there are at most two ICs in IP so the connection tables are not necessary. However, we add 
 current IC into $J$.

 \item {\bf Two degenerated ICs} (line \ref{line:two_degenerated}){\bf .} A face of $S_3$ lies in the plane of $S_2$. 
 Let $e$ be the single non-degenerate edge. We treat the two degenerate edges as faces adjacent to $e$ 
 and return them sorted like the faces given by edge faces of edge $e$.

 \item {\bf Single IC has the opposite sign to the other two} (line \ref{line:different_sign}){\bf .} 
 Let $e$ be the edge of the single IC with the different sign. 
 The plane of $S_2$ separates $e$ from the other two edges so it goes through the faces adjacent to $e$. 
 The order is determined by the function \emph{edge faces} called for the edge $e$.

 \item {\bf All ICs have the same sign} (line \ref{line:same_signs}){\bf .} Since $S_2$ is touching $S_3$ 
 at the vertex $v$, the polygon degenerates into point and thus no connection information is necessary. We add IS to $J$ and return any pair of faces.
\end{itemize}

% Either all signs are equal ($S_2$ touching $S_3$) wee , or
% a single edge have the sign oposite to other twotwo edges have a sign oposite then the thi. of According to the 
% 
% 
% and need to insert them into $I$ to correct place. To this end we exploit additional information about coincidence of points
% with faces of $S_3$. Every point $p$ from the first part of the algorithm \notePE{I would use proper references instead, I got confused, what 'part of algorithm' is..} that is not a vertex of $S_2$ belongs to a face $f$ of $S_3$; we know from the order of the points
% whether the face comes before or after the point in the intersection polygon; in the first case we set $F[0,f] =p$, in the second case we set $F[1,f]=p$. 
% Every point $p$ from the second part of the algorithm lies on an edge $e$ which is between two faces $f_0$, $f_1$. The correct order of the faces on the 
% intersection polygon is determined by the sign of the edge-triangle intersection. So, having the faces in the correct order we set $F[0,f_0]=p$ and $F[1, f_1]=p$.
\subsubsection{Ordering of intersections}
\label{sec:ordering}
The final stage of the 2d-3d intersection is ordering of ICs. If there are less then 3 ICs in the list $I$ we merge lists $I$ and $J$ remove duplicities
and get at most 2 distinguish ICs forming a degenerate IP. Otherwise, we start with first IC $p$ in $I$, put it into final list $K$, 
get its successor \nface $x = F_p(p)$ and its successor IC $q=F_q(x)$. Unless $x$ is vertex of $S_2$, we put $q$ into $K$ and repeat the process until 
we pass all ICs in $I$.

\section{Global Mesh Intersection Algorithm}
\label{sec:front_advancing}
Having the algorithms for element-element intersections at our disposal we can proceed to the mesh intersection algorithm. 
We consider the composed mesh $\mathcal T$ containing the 3d mesh $\mathcal T_1$ that we shall call a bulk mesh $\mathcal T_b$. Any other 
mesh $\mathcal T_i$, with dimensions $d_i < 3$, $i=1,\dots, N_{\mathcal T}$ we shall call a component mesh. 
We first compute all component-bulk mesh intersections that is (1d-3d and 2d-3d) using the advancing front algorithm which is described in Sections 
\ref{sec:initialization}, \ref{sec:front} and then the 1d-2d and 2d-2d 
intersections use the bulk mesh to get the intersection pair candidates, the Section \ref{sec:compnents}.

Let us consider a single pair of the component mesh $\mathcal T_c$ and the bulk mesh $\mathcal T_b$. 
Element intersections for this pair of meshes are obtained in two phases: firstly, we look for the first two elements intersecting each other (initialization); 
secondly, we prolong the intersection by investigating neighbouring elements (front tracking).
\noteJB{Not sure that "prolong" is the right verb.}

% \textbf{breadth first search} algorithm:
%   \begin{enumerate}
%     \item\label{en:first} Get next unprocessed component element $k$.
%     \item Find intersection candidates $\mathcal K$ in bulk mesh (3d elements).
%     \item \label{en:q2}For $K\in \mathcal K$ compute intersection $(k, K)$.  
%     \item Push the intersection neigbours into queues: $(k, L) \to Q_b$, $(l, K) \to Q_c$. 
%     \item While $(l, L) \in Q_b$ check intersection $(l, L)$. Append queues.
%     \item Pop pair from $Q_c$. (move to the next component element: go to \ref{en:q2}).
%     \item If $Q_c$ is empty, go to \ref{en:first}.
%   \end{enumerate}

\subsection{Initialization}
\label{sec:initialization}
\noteJB{Possibly move this afte Advancing front section since it is in fact single step in the main loop.}
We choose an arbitrary element of the component mesh and check its intersection with all elements of the bulk mesh.
This process is accelerated by computing the axis aligned bounding box (AABB) for every element and use intersection of the bounding boxes
as a fast indicator for possible intersection of the elements. This step takes $O(N)$ time with respect to the number of elements of the bulk mesh $N$.
If the number of components $k$ is small and if the components are contained inside the bulk mesh, the total time of initialization may still be linear $O(kN)$. However,
for more complex cases we organize the bounding boxes of the bulk mesh into the bounding interval hierarchy (BIH) \cite{EGWR:EGSR06:139-149},
a data structure in principle equivalent 
to R-trees \cite{guttman_r-trees:_1984}, \cite{nam_comparative_2004}. The creation time is $O(N\log(N/n))$ and the lookup time $O(\log(N/n))$ where $n$ is number
of the bulk elements in the leaf nodes of the tree.

\subsection{Advancing Front Method}
\label{front}
The advancing front algorithm requires the neigbouring information for the elements within the component mesh $\mathcal T_c$ 
as well as within the bulk mesh $\mathcal T_b$. It can be viewed as the breadth first search algorithm for a graph where graph vertices are 
intersection polygons and graph edges are sides shared by two polygons. Since every side of IP is on the boundary of either 
the component element or the bulk element, we can distinguish bulk and component edges. 
Correspondingly the candidate paris are enqueued into \emph{component queue} $Q_c$ and the \emph{bulk queue} $Q_b$, respectively.

\begin{algorithm}
  \caption{Advancing front algorithm}  
  \label{algo:advancing front}

  \SetKw{return}{return}
  \SetKw{edgefaces}{edge faces}
  \DontPrintSemicolon
  %\KwIn{vertex $v$ of $S_3$, $L[:]$ intersection results for edges of $S_3$}
  %\KwOut{$(x_1, x_2)$, $x_1, x_2 \in M_3$, coincident with $v$ and intersected by the plane of $S_2$}
  mark all $c\in \mathcal T_c$ unvisited\;
  \For{unvisited $c\in\mathcal T_c$}{
    mark $c$ visited\;
    find list $L$ of intersection candidates\;
    append $L$ to queue $Q_c$\;
    \For{$(c, b) \in Q_c$}{
        mark $c$ visited\;
        push $(c,b)$ into $Q_b$\;
        \For{$(c, b) \in Q_b$}{            
            list $I$ of ICs of $(c,b)$ intersection\;
            \For{$p\in I$}{
                \If{$p$ on boundary of $c$}{
                    get neighbours $C'$\;
                    \tcp{check pair if unvisited, mark visited}
                    enqueue $(c', b)$ into $Q_c$ for all $c'\in C'$ 
                }
                \If{$p$ on boundary of $b$}{
                    get neighbours $B'$\;
                    \If{$B'$ is empty and not on boundary of $c$}{mark $c$ unvisited}
                    enqueue $(c, b')$ into $Q_b$ for all $b'\in B'$ 
                }                
                \noteJB{Questionable}\;
                \If{$p$ on both boundaries}{
                    enqueue $(c', b')$ into $Q_c$ for all $b'\in B',\ c'\in C'$ 
                }
            }

        }
    }
  }

\end{algorithm}

%
\begin{figure}[!htb]
%   \vspace{0pt}
  \centering    
    \includegraphics[width=\textwidth]{\figpath prolongation_scheme_ip_new.pdf}
  \caption{Advancing front algorithm for 1d-2d and 2d-3d intersections.}
  \label{fig:prolongation}
\end{figure}

First, we mark all component elements $c\in\mathcal T_c$ unvisited. The for every unvisited element $c\in \mathcal T_c$, we mark it visited
then find intersection candidate pair $(c,b)$, $b\in\mathcal T_b$ 
and initialize the front tracking for every unvisited .

The key idea behind the two queues is to compute intersections for a component element with all possible bulk elements at first, and then move to a~next neighboring component element. So the bulk queue is emptied before the component queue.

On the input, we consider a candidate pair provided by the initialization (the top-left corner of the scheme).
If a nonempty intersection is found, we look for new candidate pairs among the neighboring elements
(see the white block 'Cycle over ICs').
Therefore, we iterate over the intersection corners and further exploit their topological position on the component element $c$ and the bulk element $b$.
There are 3 possible cases (applies both for 1d-3d and 2d-3d), how the intersection might be prolonged:
\begin{enumerate}[label=(\alph*)]
    \item \textbf{IC on the boundary of $c$  and inside $b$} \\
            We find all the sides $S$ coincident with the \nface of $c$ on which IC lies. Then we get all component
            elements $C'$ neighboring with $c$ over any side $s\in S$. And finally, we push all pairs $(c',b)$, $c'\in C'$ 
            into the component queue.
            Note that $c$ can have more then one neighbour component elements over the single side $s$, i.e. branches are allowed.
    \item \textbf{IC inside of $c$ anf on the boundary of $b$} \label{enum:prolong2}\\
            We find all the faces $F$ coincident with the \nface of $b$ on which IC lies. 
            Then we get all bulk elements $B'$ neighboring with $b$ over any face $f\in F$,
            analogically to the previous case.
            Finally, we push the new candidate pairs $(c, b')$, $b'\in B'$ into the bulk queue.
    \item \textbf{IC on the boundary of both $c$ and $b$} \\
            This situation includes both previously described cases. Therefore, we obtain new candidate pairs
            $(c', b)$, $c'\in C'$ and $(c, b')$, $b'\in B'$ and push them into the component queue and the bulk queue, respectively.
\end{enumerate}
% \noteJB{Can we just use two (not excluded) cases one for IB on boundary of $c$ second for CI on boundary of $b$ ?}
When the pair is taken from either of the queues, we check whether the pair has been already computed and possibly skip it. 
To this end every component element keeps a hash of the bulk elements with which we have called the intersection test.

At the end, we see that the candidate pairs are of two types: $(c, b')$ and $(c', b)$,
from which the first one goes into the bulk queue, trying to fully cover current component element 
$c$ before we proceed to the next one. We are able recognize the case when the component element 
does not lie inside the bulk elements with its full volume -- we end up in situation \ref{enum:prolong2},
but there are no more bulk neighboring elements to prolong the intersection to. Then we mark this component
element as \emph{open}, so it can be later found by initialization algorithm and possibly
covered by other bulk elements but as different component.

\subsection{Intersections between component meshes}
\label{sec:components}

% After computing the intersection of a pair of elements (line or triangle vs tetrahedron), we fill
% two queues with element pairs as candidates for further intersection. If the intersection edge 
% (point of line in 1d, edge of polygon in 2d) is inside the tetrahedron, not on its surface, we
% get a neighbouring element of the component and push it back together with the current tetrahedron into 
% \emph{component prolongation queue}. If the intersection edge is inside the \emph{slave} element 
% (line or triangle), i.e. is on the surface of tetrahedron, we get a neighbouring element of the tetrahedron
% and push it back together with the current slave element into \emph{3d prolongation queue}.

%Then we empty the two queues. 
%We pop out new candidate pairs from the \emph{bulk queue} as long as it is not empty and for every new intersection computed, 
%we repeat the previous part (means that we can further fill both queues).
%The \emph{bulk queue} is empty when the component element is fully covered by bulk elements, or
%when there is no bulk neighbor to which we can advance.
%Then we can pop a new candidate pair from \emph{component prolongation queue} and process it.
%When both queues are empty, all intersections of a component have been found and we start over by looking for the first intersection of another component.

%\notePE{We can discuss further the covering/closing of the elements and component numbering which is not tested
%thoroughly at the moment. We can show in a figure the case in 'prolong\_meshes\_13d/prolongation\_13d\_04.msh', where actually 4 components are found (therefore bulk is defined as connected 3d elements).}

% The algorithm is now unified for 1d and 2d in contrast to \cite{fris_dp_2015}, where the component prolongation
% queue is emptied at first.

% \subsubsection{1d-3d prolongation}
% 1d-3d prolongation logic
% IC is:
% - 1d element node
%     - inside tetrahedron
%         - get the neighboring 1d element
%         - push to component queue candidate pair [1d neigbor -- tetrahedron]
%     - on the surface of tetrahedron
%         - get all faces in which the IC lies (1 face, or 2 faces (edge), or 3 faces (node))
%         - get the tetrahedron neighbors on the faces
%         - push to bulk queue new candidates pairs:
%             - [current 1d element -- tetrahedron neighbor]
%             - [1d neighbor -- tetragedron neighbor]
%             - check whether the candidate pair has not been computed yet
%             (- if no new prolongation, push empty pair; means IC is on the boundary; 
%                the element is not closed then -- not fully covered with tetrahedrons)
% - inside 1d element
%     - get all faces in which the IC lies (1 face, or 2 faces (edge), or 3 faces (node))
%     - get the tetrahedron neighbors on the faces
%     - push to bulk queue candidate pair [current 1d element -- tetrahedron neighbor]
%     - check whether the candidate pair has not been computed yet
%     (- if no new prolongation, push empty pair; same meaning as above)

% A new candidate pair of elements is found during prologantion, based on the topological information of the intersection corner. There are 3 possible cases:

% \begin{itemize}
%     \item \textbf{IC lies at 1d element node and inside 3d element} \\
%             We find the neighboring 1d element over the node and push a new candidate pair [1d neighbor -- current 3d element] into component queue.
%     \item \textbf{IC lies at 1d element node and on the surface of 3d element} \\
%             We find all the faces of 3d element in which the IC lies (1 face, or 2 faces (IC on an edge), 
%             or 3 faces (IC at a node)). We find the corresponding neghboring 3d elements over the faces and
%             push the following new candidates pairs into the bulk queue: [current 1d element -- 3d neighbor], [1d neighbor -- 3d neighbor].
%             If the candidates pair has been investigated already, we skip it.
%     \item \textbf{IC lies inside 1d element (therefore must be on the surface of 3d element)} \\
%             The same as before, but we push only [current 1d element -- 3d neighbor] candidates pairs, since
%             there is no 1d neighbor.
% \end{itemize}

% 2d-3d prolongation logic CURRENT
% IC is:
% - inside 2d element side
%     - inside 3d element
%         - get the neighboring 2d elements over the side
%         - push to component queue candidates pair [2d neigbor -- current 3d element]
% - inside the face of 3d element
%     - get the neighboring 3d element over the face
%     - push to bulk queue candidate pair [2d neighbor -- 3d element]
% 
% 2d-3d prolongation logic WITH FUTURE CORRECTIONS
% IC is:
% - on 2d element side
%     - find 2d element sides (if at node) and the corresponding neighbors
%     - inside 3d element
%         - push to component queue candidates pairs [2d neigbor -- current 3d element]
%     - on the surface of 3d element
%         - get all faces in which the IC lies (1 face, or 2 faces (edge), or 3 faces (node))
%         - get the 3d neighbors over the faces
%         - push to bulk queue new candidates pairs
%             - [current 2d element -- 3d neighbor]
%             - [2d neighbor -- 3d neighbor] (all combinations)
%             - check whether the candidate pair has not been computed yet
% - inside the 2d element (=> on the surface of 3d element)
%     - get all faces in which the IC lies (1 face, or 2 faces (edge), or 3 faces (node))
%     - get the 3d neighbors over the faces
%     - push to bulk queue new candidates pairs [current 2d element -- 3d neighbor]
%     - check whether the candidate pair has not been computed yet

\section{Benchmarks}
\label{sec:benchmarks}

%done \noteJB{I prefer to call them element intersection. Fundamental intersection seems to epic.}
\noteJB{We compare just against NGH which we should briefly describe. Neither of the algorithms is optimized. The new one deals correctly with special cases and provides 
barycentric coordinates (so more work for less time).}
\noteJB{Also write little about Flow123d here. The language etc.}

In this section, we present numerical results on several benchmark problems. 
In the beginning, we provide some software development related information.
Then we compare the performance of our algorithms for element intersections with other approaches \noteJB{they are all ours}.
Next we shall compare our algorithms with different initialization phase (candidate pairs search),
and using the advancing front method or not. We shall show the results both on a mesh of a real locality
and an artificial mesh.

\subsection{Software Flow123d}
The implementation of the presented algorithms is part of the open source software Flow123d~\cite{flow123d}.
The C++ is used as principal language. Flow123d provides models for saturated and unsaturated groundwater flow,
soulute transport with sorption, and  heat transfer. Mixed finite elements, finite volumes, and discontinuous Galerkin method are 
used. All equations are consistently formulated for meshes of mixed dimensions.  However,
non-matching meshes are at the moment available only in experimental 1d-2d flow model, using mortar like coupling.
% This work is the first greater step on the way to dimensional coupling models on non-matching meshes 
% in Flow123d.
Part of the software is also a legacy code for element intersections that we refer to as \ngh.

% \notePE{check NGH.. how to refer to NGH later?}
% Currently, there is also another implementation of the element intersections in the software (we shall refer to
% it later as \ngh). \noteJB{Desribe later different ways how to compute line-tringle intersection}. 
% It uses standard means of analytic geometry for computation of intersections line-line and line-plane.
% This leads to solving a large amount of small linear systems of equations. \noteJB{And computing orientation determinants}
% This algorithm also does not provide us barycentric coordinates and topological position of ICs. \noteJB{Although it could be done in principle.}
% At last, the algorithm cannot handle all of the special cases.
% We must note, that neither the old algorithms nor the new ones are particularly optimised at the moment.

%Note, that we use Armadillo~\cite{armadillo_2016}, a C++ linear algebra library, for linear algebra operations. 
%Most importantly, all coordinates are represented as Armadillo vectors, so we can use dot and cross product and 
%further vector and matrix operations provided by the library.

\subsection{Theoretical comparison}
Before presenting the results of our benchmarks, let us try to estimate the number of arithmetic operations performed during 1d-3d and 2d-3d element intersections
as we use different algorithms for the elementary line-triangle intersection. We consider 4 algorithms for the line-triangle intersections: \plucker algorithm 
described in Section \ref{sec:1d-2d}, the legacy code \ngh, the plane clipping algorithm due to Haines \cite{haines_fast_1991}, the minimum storage
algorithm due to Möller and Trumbore (MT) \cite{moller_fast_1997}. Raw numbers of arithemtic operations are summarized in Table \ref{tab:fundamental_flops}.
For the \plucker and \ngh we count FLOPS from the implementation of individual intersection algorithms, for Haines and MT we estimated FLOPS in theoretical 
implementation. In particular we account for reuse of the calculations in \plucker and Haines algorithms. Conclusions from this census are:
algorithms based on the \plucker coordinates should be competitive with state of art algorithms in the case of 1d-3d and 2d-3d intersections. The expected performance
for the 1d-2d case seems to be poor however these intersections are computed after 1d-3d and 2d-3d so the \plucker coordinates may be reused, then we get 
quite competitive $45$ FLOPS. Similarly we may expect better times in remaining two intersection cases when we account for the reuse of calculation between elements.
Compared to the legacy code \ngh the \plucker algorithm for 1d-3d should be about $2.5\times$ faster and the 2d-3d about $5\times$ faster.

\noteJB{Need better timing for \plucker with respect to NGH in 
2d-3d  case. Could it be due to polygon tracing?}

\begin{table}[!htb]
    \caption{Raw number of FLOPS used by different intersection algorithms.}    
    \label{tab:fundamental_flops}
    \centering
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}[b]{l|lll}
            algorithm           & 1d-2d         & 1d-3d         & 2d-3d \\                      
            \hline\\
            \plucker           & 92            & 198           & 426\\
            Haines              & 51            & 177           & 469\\
            Moller and Trumbore & 42            & 168           & 756\\
            \ngh                & 119           & 476           & 2142\\
    \end{tabular}

\end{table}

\subsection{Element Intersection Algorithms}
The first benchmark focuses on the 1d-3d and 2d-3d element intersections. We randomly generated
100000 element pairs inside a unit cube, from which approximately 65\% have nonempty intersection.
In some of the empty cases, the bounding boxes did not collide, so the actual intersection algorithms were skipped. 
A single element pair was computed 100 times to obtain reasonable computational time.

% \vspace{2ex}
% \begin{minipage}{\textwidth}
%     \centering
%     \hspace{-30pt}
%   \begin{minipage}[b]{0.44\textwidth}
%     \centering
%     \renewcommand{\arraystretch}{1.2}
%     \begin{tabular}[b]{l|l}
%             algorithm, 2d-3d& FLOPs\\
%                       & estimate\\
%             \hline\\
%             parametric plane & 585\\
%             normal plane (reuse) & 765\\
%             Moller and Trumbore & 783\\
%             Pl\"ucker (reuse) & 306
%     \end{tabular}
%     \captionof{table}{Estimation of floating point operations count and comparison of different approaches.
%                       Only 2d-3d case considered.}
%     \label{tab:fundamental_flops}
%     \notePE{comment on different strategies}\noteJB{Rather put into text. I shall comment that.}
%   \end{minipage}
%   \hfil
%   \begin{minipage}[b]{0.5\textwidth}
%     \centering
%     \includegraphics[width=\textwidth]{\figpath intersections_fundamental_speed.pdf}
%     
%     \captionof{figure}{Efficiency of \plucker approach compared to the legacy \ngh code.}
%     \label{fig:fundamental_speed}
%     \end{minipage}
% \end{minipage}
% \vspace{2ex}

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\textwidth]{\figpath intersections_fundamental_speed.pdf}
    \caption{Efficiency of \plucker approach compared to the legacy \ngh code.}
    \label{fig:fundamental_speed}
\end{figure}

We see the benchmark results in the \fig{fig:fundamental_speed}. Green values correspond to the presented algorithms
using \plucker coordinates, red values to \ngh code. The gained speed up factor is approximately 5 in 1d-3d case and 2 in 2d-3d. Further we provide estimated count of floating point operations for different approaches in the Table \ref{tab:fundamental_flops}.
\notePE{Improve and complete...}

\subsection{General algorithms}
In previous sections, we have suggested an intersection algorithm that uses axes aligned bounding boxes, BIH to find initial
intersection and advancing front method (shortly AFront) to prolong the intersection. From now on, we shall refer to the algorithm made of these components \emph{BIHsearch}. In this section, we shall compare two additional algorithms \emph{BBsearch} and \emph{BIHonly} to see the effects of BIH and AFront (see Table \ref{tab:search_algorithms}). \emph{BBsearch} does not use BIH, but searches for the initial intersection using only AABB, and then follows AFront. The version 
\emph{BIHonly} computes AABB together with BIH, but does not use AFront.

\begin{table}[!htb]
    \centering
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{>{\centering\arraybackslash}p{20ex}
                    >{\centering\arraybackslash}p{20ex}
                    >{\centering\arraybackslash}p{20ex}}
        \colorbox{NavyBlue}{\textcolor{white}{BIHsearch\vphantom{By}}} & \colorbox{orange}{BBsearch\vphantom{By}} & \colorbox{Goldenrod}{BIHonly\vphantom{By}} \\ \hline
        BIH(AABB) & AABB & BIH(AABB) \\
        AFront & AFront & ---
    \end{tabular}
    \caption{Structure of the general algorithms. 'AFront' stands for the proposed advancing front method.
    \noteJB{Just in the text. Possibly use different name for variants.}
    \notePE{jsem vzal z posledni prezentace.. prislo mi to prehledny.. vim jak jsem se v diplomce VF hrozne krkolome vracel od grafu k popisu algoritmu..}}
    \label{tab:search_algorithms}
\end{table}

Let us now start with the problem on an artificial mesh.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.43\textwidth]{\figpath test1_cube0.pdf}
    \caption{Artificial mesh -- a cube with two perpendicular planes placed on the diagonals of the cube.
             The planes are also nonmatching, therefore can be seen as two independent components.}
    \label{fig:cube_mesh}
\end{figure}

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.8\textwidth]{\figpath intersections_test1_cube_speed_scaling.pdf}
    \caption{Dependence of the algorithms performance on number of intersection (and mesh refinement). }
    \label{fig:cube_speed}
\end{figure}

Next, we study the performance of the intersection algorithms on a mesh of a real problem, see \fig{fig:bedrichov_meshes}. The mesh represents a mountain ridge in the Jizera mountains which 
includes a system of geological fractures (\fig{fig:bedrichov_cut}). We also used this model to create
a less real mesh, in which the fractures extend through the bulk surface (\fig{fig:bedrichov_extend}). 
However, this situation might rather occur in another application than aporous media problem.
In total, 28 fractures are in the mesh.

The results for both meshes can be seen in the \fig{fig:bedrichov_speed}, pay attention to the different time
scales in the graph. In the first case, we notice that algorithms using AFront are nearly twice as fast as \emph{BIHonly}. The speed up corresponds to the time saved by using AFront instead of searching the whole mesh for candidate pairs. Computation of BIH in \emph{BIHsearch} pays off and the algorithm performs better than \emph{BBsearch}.

In the second case, we observe large blow up for \emph{BBsearch}. It is caused by the exterior elements, 
for which all bulk elements bounding boxes are iterated, so we can be sure there is no intersection.
This of course is much better performed using BIH.

In contrast to the artificial case on the fractured cube, BIH construction is worth due
to the presence of more components. Having more fractures inside the bulk would result to more dramatic difference.

\begin{figure}[!htb]
    \subfloat[interior fractures]{
        \includegraphics[width=0.49\textwidth]{\figpath bedrichov_fract.pdf}
        \label{fig:bedrichov_cut}
    }
    \hspace{3pt}
    \subfloat[extending fractures]{
        \includegraphics[width=0.49\textwidth]{\figpath bedrichov_fract_extend.pdf}
        \label{fig:bedrichov_extend}
    }
    \caption{A mesh of the real locality of Bed{\v r}ichov in the Jizera mountains.
             We see fractures inside the bulk mesh in the left figure,
             fractures are extending the bulk mesh.}
    \label{fig:bedrichov_meshes}
\end{figure}

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.8\textwidth]{\figpath intersections_bedrichov_both_speed.pdf}
    \caption{Comparison of the algorithms on meshes of Bed{\v r}ichov locality -- interior fractures on the left,
             extending fractures on the right.}
    \label{fig:bedrichov_speed}
\end{figure}

\section{Conclusions}
\label{sec:conclusins}

TODO:
- line intersection tracking for accelerate 2d-2d intersections
- better handling of special cases in paarticular in relation to prolongations
- better calculation reuse (pass with prolongations)
- optimisation of element intersection - skip unnecessary calculations
%\section{Acknowledgement}
%The paper was supported in part by the Project OP
%VaVpI Centre for Nanomaterials, Advanced Technologies  and Innovations
%CZ.1.05/2.1.00/01.0005.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%%  \bibliographystyle{elsarticle-harv} 
%%  \bibliography{<your bibdatabase>}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

% \begin{thebibliography}{00}
% 
% %% \bibitem[Author(year)]{label}
% %% Text of bibliographic item
% 
% \bibitem[ ()]{}
% 
% \end{thebibliography}
 %\nocite{dip}
 %\bibliographystyle{elsarticle-harv} 
 %\bibliographystyle{elsarticle-num-names} 
 \bibliographystyle{elsarticle-num} 
 \bibliography{paper_intersections.bib}
\end{document}

% \cite{sukumar_extended_2000} - Use {\emph geometric predicates} incircle and orientation. Use mesh of bounding boxes to search for intersection candidates. 
% Only use point queries for all vertices of triangle/tetrahedron. Use compatible partitioning of elements by the crack elements.
% 
% \cite{boon_robust_2016} Robust discretization of porous media. Complete error analysis with mortar non-compatible case. Seems they use normal fluxes accros
% fracture instead of pressure traces. This allows conforming discretization and error results.
% 
% \cite{reichenberger_mixed-dimensional_2006} Finite volume method, fully implicit two phase flow. Cite usage of FRAC3D generator. 
% 
% \cite{sladkova_bakalarka} Space-time DG with adaptive mesh, need integration of product of functions on different meshes (in 2d). An algorithm 
% for computing triangle-triangle intersections in 2d.
% 
% \cite{pichot_generalized_2012} DFN + MHFEM. Mortar method for intersections. Focused on formulation not on intersections. 
% 
% \cite{flemisch_non-matching_2012} (Barbara) Mortad, non=matching, acustics. Theory, structured gird case. No complex intersections.
% 
% \cite{mustapha_hybrid_2014} Patent for curiosity.
% 
% \cite{mittal_immersed_2005} Immersed boundary. ??
% 
% \cite{owen_survey_1998} ??
% 
% \cite{kuznetsov_overlapping_1998}
% 
% \cite{platis_fast_2003} Original algorithm for line-thetrahedra intersections.
% 
% \cite{haines_fast_1991} Line-convex Polyhedra intersections using generalisation of line-clipping algorithm
% 
% 
% USED:
% \cite{massing_efficient_2013} - Implementation of the Nitche method in Fenics. They need 2d-3d intersections, but only for distribution od quadrature points. 
% So they do not store whole intersection data, but only the quad points. The boundary of one domain is partitioned into intersections with elements of the other domain.
% Use {\\emph collision relations} ( pairs of faces of boundary and elmements of the background mesh) and {\emph collision maps} 
% ( map face to all intersections and background element to all intersections, can be constructed from the collision relation), They cite books from computer graphics ..
% Mention geometric predicates. Present algorithms for:  find intersection candidates, comput intersections, integrate over complex domains.
% 
% \cite{elsheikh_reliable_2012} Claim to provide robust algorithm covering all tiangle-trinagle degenerate cases. Basic idea is the same as our: use some tree 
% (Binary SPace Tree) structure to find initial point on intersection curve and trace the curve in both directions. Quite clever data structure for the mesh. Adaptive precision 
% geometric predicates. Store topological information about the intersection corners (point on triangle, edge, vertex). Resolve various degenerate cases, 
% dicuss mesh optimization after subdivision.
% 
% 
% \cite{gander_algorithm_2013} Full algorithm for 2d=3d called PANG. Implemented in DUNE in 2010 (Bastian). Other software (Hecht, MpCCI). 
% Triangle-Triangle and Tetra-Triangle intersections done similarly as in aur approach: set of points including vertices and then sort them countercolckwise.
% Local intersection algorithm do not provide neighbouring information so the traversal algorithm do not use it.
% 
% Review of algorithms: \cite{gander_algorithm_2009}
% 
% 
% \cite{bournival_mesh-geometry_2008} Application of mixed-dimensional approach in mechanics. Beam and shell elements ...

%\endinput
%%
%% End of file `elsarticle-template-harv.tex'.

-